Changing Flights In The Air\\ (Abortable Chain Agreement)

### 2015-05-14 paper/intro.tex
Work on the introduction.

However, devising new SMR algorithms, optimized with particular situations or metrics in mind, is hard: today, several decades after the first SMR algorithms \cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion} were published, devising new SMR algorithms is still hard enough to justify a publication in a top conference \cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.

@@

We propose a framework to make existing SMR algorithms modular and add new optimizations incrementally, without modification to the existing. 
An algorithm in our framework is composed of independent modules that can run, abort their execution, and pass the baton to any other module.
Our framework has the following features:
\begin{itemize}
  \item Effortless re-use of existing modules: A SMR algorithm expressed in our framework can can be analyzed, tested, and proved correct in isolation and once and for all.  Our composition theorem guarantees that the composition of any number of correct modules is a correct SMR algorithm. 
  \item Simplified algorithms thanks to a clear separation of concerns: one can devise a new specialized SMR algorithm focusing only on the situation for which it is specialized.  A composable SMR algorithm is not required to make progress in all situations: if it does not, the system can switch to another SMR algorithm.

  \item Seamless combination of different SMR algorithms: one can combine different SMR algorithms, optimized for different situations and metrics, without any changes and with guaranteed correctness.
  \item Practical
    \begin{itemize}
  \item Tolerating $k$ fault with $k+1$ replicas and a reliable reconfiguration master. 
  \item Built-in support for crucial practical optimizations such as executing commuting commands without synchronization and expediting the execution of read-only commands.
    \end{itemize}
\end{itemize}

A composite SMR algorithm built with our framework works by running the SMR module deemed the most appropriate at any given time, dynamically changing SMR module when needed.
Thus algorithms designers who want to create a new SMR algorithm, optimized for particular conditions, can focus only on the particular conditions of interest and reuse existing modules to handle the other cases. All the performance analysis, tests, and proofs of the existing SMR modules remain valid and only the new SMR needs to be analyzed.
Our framework thus enables incremental development of SMR algorithms, where each new development can build on the previous ones instead of starting again from scratch.

An important practical aspect of SMR is fast reconfiguration to replace faulty components, without which the stability of a system is inevitably compromised.

To be usable in  practice, SMR implementations must have high performance, be it throughput, latency, or some other metric depending on the setting, and resort to many optimizations which a framework like ours must support to be practical.
Two key practical aspects of SMR are the ability to reconfigure the system to replace faulty components and the ability to process commutative commands without synchronization.
There lies our key contribution: 
Our framework is based on Generalized Consensus, a formulation of the State-Machine Replication problem that allows processing commutative commands without synchronization.

At the heart of our framework lies a simple abstraction of the familiar notion of rounds (also called epochs, ballots, or views in the literature) used in most SMR algorithms. To illustrate our approach, let us consider Fast Paxos, one of the first examples of SMR algorithm which can change mode dynamically.
Fast Paxos is a round-based SMR algorithm that has two types of rounds: fast rounds and normal rounds. Fast rounds process requests faster than normal rounds in contention-free executions but fall back to a slower scheme under contention. Normal rounds are the same as in the classic Paxos. Any process can initiate a round change at any point during execution: this allows switching between fast and normal rounds as well as, similarly to classic Paxos, changing the leader from one normal round to
another. Each kind of round maintains different invariants relating the state of each process and the the values that might have been chosen in the previous rounds, thus a normal round must know which past rounds are fast or normal rounds in order to take over the execution safely. When adding more kinds of rounds such an approach will quickly become unmanageable, as each type of rounds must have one case for each possible other kind of round, making $n^2$ cases to consider in total.

Our notion of Composable Generalized Consensus defines a common interface and a common correctness property for a round, abstracting over the particular round implementations and making all rounds compatible with each other. 
Moreover, our interface for rounds enable a straightforward crucial optimization seen in practical systems: introducing an external reconfiguration master which manages the addition of new replicas and the removal of failed ones. Interestingly, as explained in \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, in a real system it is often possible to implement a reliable reconfiguration master which allows to tolerate $f$ faults with $f+1$ replicas instead of the traditional $2f+1$ replicas.

The reconfiguration master can also make the choice of the new type of round to execute, providing the system's reconfiguration policy.

>>>>

However, devising new SMR algorithms, optimized with particular situations or metrics in mind, is hard: today, more than two decades after the first SMR algorithms \cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion} were published, devising new SMR algorithms is still hard enough to justify a publication in a top conference \cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.

@@

We propose a framework to make existing SMR algorithms modular and add new optimizations incrementally, without modification to the existing. 
An algorithm in our framework is composed of independent modules that can run, abort their execution, and pass the baton to any other module.

Crucially, when adding a new module to an existing algorithm, all the performance analysis, tests, and proofs of the existing modules remain valid and only the new SMR module needs to be analyzed: Our composition theorem guarantees that the composition of any number of correct modules is a correct SMR algorithm, thus one can combine different SMR algorithms, optimized for different situations and metrics, without any changes and with guaranteed correctness.
Our framework therefore enables incremental development of SMR algorithms, where each new development can build on the previous ones instead of starting again from scratch.

Moreover, our framework enables a clear separation of concerns  which simplifies the designing individual modules.
A composite SMR algorithm built with our framework works by running the SMR module deemed the most appropriate at any given time, dynamically changing SMR module when needed. Therefore a composable SMR algorithm is not required to make progress in all situations: if it does not, the system can switch to another SMR algorithm.
Thus algorithms designers who want to create a new SMR algorithm, optimized for particular conditions, can focus only on the particular conditions of interest, knowing that the other existing modules will handle the other cases.

The key novelty of our framework lies in its practical applicability. In practice, a SMR algorithm must support fast reconfiguration to replace faulty components, without which the long-term stability of a system is inevitably compromised. Our framework clearly separates the execution of commands from the reconfiguration decisions (including changing SMR module), allowing the use of an external reconfiguration master. This enables tolerating $k$ failures with $k+1$ replicas, a very
important improvement over the traditional $2k+1$ needed replicas.
Moreover, in practice, commands sent to a SMR algorithm often commute. Our framework is based on Generalized Consensus, a formulation of the State-Machine Replication problem that allows processing commutative commands without synchronization.

TODO: expand on the preceding paragraph.

Our framework is fully formalized and its composition theorem proved in Isabelle/HOL. Moreover, to quickly check concrete modules, we also provide a TLA+ specification of our framework which allows model-checking algorithms with the TLC model-checker.

TODO: the two abstract specs.

At the heart of our framework lies a simple abstraction of the familiar notion of rounds (also called epochs, ballots, or views in the literature) used in most SMR algorithms. To illustrate our approach, let us consider Fast Paxos, one of the first examples of SMR algorithm which can change mode dynamically.
Fast Paxos is a round-based SMR algorithm that has two types of rounds: fast rounds and normal rounds. Fast rounds process requests faster than normal rounds in contention-free executions but fall back to a slower scheme under contention. Normal rounds are the same as in the classic Paxos. Any process can initiate a round change at any point during execution: this allows switching between fast and normal rounds as well as, similarly to classic Paxos, changing the leader from one normal round to
another. Each kind of round maintains different invariants relating the state of each process and the values that might have been chosen in the previous rounds, thus a normal round must know which past rounds are fast or normal rounds in order to take over the execution safely. When adding more kinds of rounds such an approach will quickly become unmanageable, as each type of rounds must have one case for each possible other kind of round, making $n^2$ cases to consider in total.

Our notion of Composable Generalized Consensus defines a common interface and a common correctness property for a round, abstracting over the particular round implementations and making all rounds compatible with each other. 
Moreover, our interface for rounds enable a crucial optimization seen in practical systems: introducing an external reconfiguration master which manages the addition of new replicas and the removal of failed ones. Interestingly, as explained in \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, in a real system it is often possible to implement a reliable reconfiguration master which allows to tolerate $f$ faults with $f+1$ replicas instead of the traditional $2f+1$ replicas.
The reconfiguration master can also make the choice of the new type of round to execute, providing the system's reconfiguration policy.
